Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees

### 2023-05-27 script/manuscript_MEE.Rmd
JBF ST edits throughout intro added

Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogeneization (REF), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity). Monitoring ecological communities structure and interaction will be critical to informing and prioritizing conservation efforts (REF), particularly considering the large proportion of know species currently threaten or likely to be threaten in the near future (@pimm2014biodiversity; @pereira2010scenarios). However, ecosystem monitoring remains inconsistent in time and space (@yoccoz2001monitoring) as it requires substantial resources, thereby leaving several regions, ecosystems, and even species under-observed (@pereira2012global; @collen2008tropical; @meyer2015global; @ruete2015displaying).

@@

  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), within a framework which utilizes custom species sequence databases derived via species distribution modelling, and temporal filtering. To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific sequence database relevant to the region of study and the ecological characteristics of interest (@bell2022plants). To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g. iNaturalist). We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area. This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers. This approach can significantly reduce processing time, increasing the a projects efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.

@@

  Plant flowering varies over space and also in time, and in particular, contrasts in the flowering periods of many plant species can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC, @thompson1994coevolutionary, @caradonna2021seeing). In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction, @caradonna2020temporal). We work to develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new, Figure 1).

```{r Filter Diagram, out.width = "35%", fig.cap="Simplified Conceptual Diagram of three approaches leading towards classification of sequencing results, and the number of species associated with them in our area. The upper three boxes indicate a common coarse approach, assuming one has a digitizewd Flora, which is not always the case. The center two boxes indicate the computational approach illustrated here. The final box indicates the use of the expert field data in the case study. The stem of the final applies to both Computational and Expert Survey results, and should be thought of as using time ala chromatography, in the post-classification process."}
knitr::include_graphics('../graphics/plots/filter_diagram.png')
```

  We tested this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study. In particular, we use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence). The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne). The two foraging phases of the queen bumblebee life cycle are essential to 1) increase their weight before diapause, and 2) increase their ovary weights while establishing their recently found nests. Both time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble). Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline). These areas often represent the most diverse areas in the temperate and often offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate). An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be incorporated into their current diversity (@loarie2009velocity, @dobrowski2016climate).

@@

>>>>

Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogenization (@olden2004ecological, @wang2021biotic), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity). Monitoring ecological communities structure and interaction will be critical to informing and prioritizing conservation efforts (REF), particularly considering the large proportion of know species currently threaten or likely to be threaten in the near future (@pimm2014biodiversity; @pereira2010scenarios). However, ecosystem monitoring remains inconsistent in time and space (@yoccoz2001monitoring) as it requires substantial resources, thereby leaving several regions, ecosystems, and even species under-observed (@pereira2012global; @collen2008tropical; @meyer2015global; @ruete2015displaying).

@@

  To address these challenges, we are proposing the generation of a smaller, and more appropriate reference list for identifying candidate taxa from barcode data. By reducing the list of likely species in a study site it can increase the accuracy and efficiency of metabarcoding results in plants, which will better inform data collection for projects with limited access to taxonomic expertise. Specifically, we propose using current distribution data to generate a list of candidate species relevant to the study area and its ecological characteristics (@bell2022plants). Initially such a species list can be leveraged to identify a target collections list of plant tissues to be included in genetic studies. It can subsequently help to reduce the size of a reference sequence database, which increases computation speed and efficiency, allowing for the use of genomic data on personal computers. This approach can significantly reduce processing time, increasing  project efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.  We test this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study. We use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence). The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen, @johnson2021airborne).

@@

```{r Filter Diagram, out.width = "35%", fig.cap="Simplified Conceptual Diagram of three approaches leading towards classification of sequencing results, and the number of species associated with them in our area. The upper three boxes indicate a common coarse approach, assuming one has a digitized Flora, which is not always the case. The center two boxes indicate the computational approach illustrated here. The final box indicates the use of the expert field data in the case study. The stem of the final applies to both Computational and Expert Survey results, and should be thought of as using time ala chromatography, in the post-classification process."}
knitr::include_graphics('../graphics/plots/filter_diagram.png')
```

@@

To improve metabarcoding reliability and efficiency, we suggest creating a regional list of candidate species using digital collections gleaned from herbaria, survey work, and community science  (Figure 1). This list can further be refined using species distribution models and temporal filtering to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area. The final list is then used to inform collection of plant samples to create a library and inform metabarcoding . We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations.
